### 2015-05-14 paper.tex
Minor edits

\subsection{k-medians}
The algorithm we will focus on is the k-medians algorithm. The k-medians algorithm assumes a space $(X, d)$ for which $X$ is the set of possible points and distance $d : ({ X \choose 2}) \rightarrow R_{\geq 0}$ which satisfies the triangle inequality. We are then given a set of points $S \subset X$. The goal is to label the points in $S$ each into one of k clusters $C_1, C_2, ..., C_k$ with corresponding representative points $c_1, c_2, ..., c_k \in X$. This defines a clustering $C$. The optimal objective clustering is that which minimizes $\Phi(C) = \sum\limits_{i = 1}^k \sum_{x \in C_i} d(x, c_i)$.

@@

The goal of this work is to develop a polynomial runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above. It is known that a better than $1 + \frac{2}{\delta}$ approximation for k-medians is NP-hard \cite{JMS02}. In this review we will show explore the following results:

@@

\item Assume that the (k-1)-medians optimal cluster is more expensive than the k optimal clustering by a factor of $(1+\alpha)$ where $\alpha > 0$. Then it is possible to achieve a ($1+\epsilon$) approximation to the optimal k-medians in time polynomial in $n$ and $k$, and exponential in $1/\epsilon$ and $1/\alpha$, where $n = |S|$.

@@

\begin{algorithm}
\caption{Cluster TODO: better name}\label{alg:cluster}
\begin{algorithmic}[1]
\Procedure{Cluster}{}
\BState \textbf{Initialization Stage:} $Q \gets Q_{init}$
\BState \textbf{Population Stage:} For $s=n, n-1, n-2, \ldots 1$ do:
\State a) Set $r=\frac{\beta OPT}{4 s}$
\State b) Remove any point $x$ such that $d(x, Q) < 2r$.
\State c) For any remaining data point $x$, denote the set of data points whose distance from $x$ is at most $r$ by $B(x, r)$.  Connect any two remaining points $a$ and $b$ if: (i) $d(a, b) \lq r$ (ii) $|B(a, r)| > \frac{s}{2}$ and (iii) $|B(b, r)| > \frac{s}{2}$.
\State d) Let $T$ be a connected component of size $ > \frac{s}{2}$.  Then i) Add $T$ to $Q$.  ii) Define the set $B(T) = \{x : d(x, y) \leq 2r$ for some $y \in T\}$.  Remove the points $B(T)$ from the instance.
\BState \textbf{Centers-Retrieving Stage}: For any choice of $k$ components $T_1, T_2, \ldots, T_k$ out of $Q$:
\State a) Find the best center $C_i$ for $T_i \cup B(T_i)$.
\State b) Partition all $n$ points according to the nearest point among the $k$ centers of the current $k$ components.
\State c) If a clustering of at most $(1 + \epsilon)OPT$ is found - output these $k$ centers and halt.
 %\hspace{\algorithmicindent} $pq.put( (v_j, e_{jk}), e_{jk}^z)$.
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

We know all points in $T$ are within $\frac{\beta\text{OPT}}{2|C^*_i|}$ of $c_i^*$ because at least some are in the inner ring of $C^*_i$. Following from \ref{lemma:4.2a} and \ref{lemma:4.2b}, it is then impossible for a point to exist that is further than $\frac{\beta\text{OPT}}{2|C^*_i|}$ from $c_i^*$. This immediately implies that $T \subset C_i^*$ because $C_i^*$ is beta distributed. Additionally we have that all points in B(T) are less than $2r$ away from some point in $T$. Because we established $2r = \frac{\beta\text{OPT}}{4s} \leq \frac{\beta\text{OPT}}{2|C_i^*|}$, all points in B(T) then within $ \frac{\beta\text{OPT}}{|C_i^*|}$ of $c_i^*$, and therefor within $C_i^*$.

Because all points in $T$ are sufficiently close to $C_i^*$ there are at most $|C_i^*|$ points in $T$, and so $s < 2|C_i^*|$. All inner ring points are at most $\frac{\beta\text{OPT}}{4|C^*_i|} < 2r$ away from any other point, requiring all points not in $T$ be added to $B(T)$. Therefor no other component can contain inner ring points for $C_i^*$.

@@

\item $C^*_i$ is a cheaper cluster and $s \geq |C^*|$. Because we established $d(c^*, y) < \frac{\beta \text{OPT}}{2|C^*|}$ or $d(c^*, y) > \frac{3 \beta \text{OPT}}{4|C^*|}$ We know at least one point must meet the larger property. Since the inner ring contains $> |C^*|/2$, but $T$ also contains at least $s/2 = |C^*_i|/2$ points, some point in $T$ must not be in $C^*_i$, and therefor have distance larger than $\frac{3 \beta \text{OPT}}{4|C^*|}$. \ref{lemma:4.2b} then tells us all points must meet the property.

@@

We will show that $d(c_i, c_i*) < \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$ for a cheap cluster. We know by the definition of a cheap cluster that the cost is at most $(\epsilon/32)\beta\text{OPT}/|C_i^*|$. In order not exceed this bound, at least half of the points have distance at most $(\epsilon/16)\beta\text{OPT}/|C_i^*|$ from $c_i^*$. Also these are by definition in the inner ring. We know that at least half the points in a cheap cluster reside in the inner ring. If $d(c_i, c_i^*)  \geq \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$, then we require at least half of the points to contribute at least $(\epsilon/16)\beta\text{OPT}/|C_i^*|$ to the sum, which means some point in the inner ring must do so. Therefor $\sum\limits_{x\in T\cup B(T)}d(x, c_i)> OPT_i$, which contradicts the optimality of the inner ring points.

We therefor for each $x \in C_i^* and x\notin T\cup B(T)$ we have $d(c_i, c_i^*) \leq \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$. So $d(x, c_i) \leq d(x, c_i^*) + d(c_i^*, c_i) \leq (1+\epsilon)d(x,c_i^*)$. And so each point adds at most a factor of $(1+\epsilon)$ to the error, bounding our total error to $(1+\epsilon)\text{OPT}$

>>>>

\subsection{k-medians}
The algorithm we will focus on is the k-medians algorithm. The k-medians algorithm assumes a space $(X, d)$ for which $X$ is the set of possible points and distance $d : ({ X \choose 2}) \rightarrow R_{\geq 0}$ which satisfies the triangle inequality. We are then given a set of points $S \subset X$. The goal is to label the points in $S$ each into one of k clusters $C_1, C_2, ..., C_k$ with corresponding representative points $c_1, c_2, ..., c_k \in X$. This defines a clustering $C$. The optimal objective clustering is that which minimizes $\Phi(C) = \sum\limits_{i = 1}^k \sum_{x \in C_i} d(x, c_i)$.  For $k$-medians, $d$ is the $L_1$ norm.

@@

The goal of these papers is to develop a polynomial runtime clustering algorithm with minimal constraints on the values of $c$ and $\delta$ as described above. It is known that a better than $1 + \frac{2}{\delta}$ approximation for k-medians is NP-hard \cite{JMS02}. In this review we will show explore the following results:

@@

\item Assume that the ($k-1$)-medians optimal cluster is more expensive than the $k$ optimal clustering by a factor of $(1+\alpha)$ where $\alpha > 0$. Then it is possible to achieve a ($1+\epsilon$) approximation to the optimal k-medians in time polynomial in $n$ and $k$, and exponential in $1/\epsilon$ and $1/\alpha$, where $n = |S|$.

@@

\begin{algorithm}
\caption{$k$-Median Approximation Clustering (as in \cite{secondpaper})}\label{alg:cluster}
\begin{algorithmic}[1]
\Procedure{Cluster}{}
\BState \textbf{Initialization Stage:} $Q \gets Q_{init}$
\BState \textbf{Population Stage:} For $s=n, n-1, n-2, \ldots 1$ do:
\State a) Set $r=\frac{\beta OPT}{4 s}$
\State b) Remove any point $x$ such that $d(x, Q) < 2r$.
\State c) For any remaining data point $x$, denote the set of data points whose distance from $x$ is at most $r$ by $B(x, r)$.  Connect any two remaining points $a$ and $b$ if: (i) $d(a, b) \lq r$ (ii) $|B(a, r)| > \frac{s}{2}$ and (iii) $|B(b, r)| > \frac{s}{2}$.
\State d) Let $T$ be a connected component of size $ > \frac{s}{2}$.  Then i) Add $T$ to $Q$.  ii) Define the set $B(T) = \{x : d(x, y) \leq 2r$ for some $y \in T\}$.  Remove the points $B(T)$ from the instance.
\BState \textbf{Centers-Retrieving Stage}: For any choice of $k$ components $T_1, T_2, \ldots, T_k$ out of $Q$:
\State a) Find the best center $C_i$ for $T_i \cup B(T_i)$.
\State b) Partition all $n$ points according to the nearest point among the $k$ centers of the current $k$ components.
\State c) If a clustering of at most $(1 + \epsilon)OPT$ is found - output these $k$ centers and halt.
 %\hspace{\algorithmicindent} $pq.put( (v_j, e_{jk}), e_{jk}^z)$.
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

We know all points in $T$ are within $\frac{\beta\text{OPT}}{2|C^*_i|}$ of $c_i^*$ because at least some are in the inner ring of $C^*_i$. Following from \ref{lemma:4.2a} and \ref{lemma:4.2b}, it is then impossible for a point to exist that is further than $\frac{\beta\text{OPT}}{2|C^*_i|}$ from $c_i^*$. This immediately implies that $T \subset C_i^*$ because $C_i^*$ is beta distributed. Additionally we have that all points in B(T) are less than $2r$ away from some point in $T$. Because we established $2r = \frac{\beta\text{OPT}}{4s} \leq \frac{\beta\text{OPT}}{2|C_i^*|}$, all points in B(T) then within $ \frac{\beta\text{OPT}}{|C_i^*|}$ of $c_i^*$, and therefore within $C_i^*$.

Because all points in $T$ are sufficiently close to $C_i^*$ there are at most $|C_i^*|$ points in $T$, and so $s < 2|C_i^*|$. All inner ring points are at most $\frac{\beta\text{OPT}}{4|C^*_i|} < 2r$ away from any other point, requiring all points not in $T$ be added to $B(T)$. Therefore no other component can contain inner ring points for $C_i^*$.

@@

\item $C^*_i$ is a cheaper cluster and $s \geq |C^*|$. Because we established $d(c^*, y) < \frac{\beta \text{OPT}}{2|C^*|}$ or $d(c^*, y) > \frac{3 \beta \text{OPT}}{4|C^*|}$ We know at least one point must meet the larger property. Since the inner ring contains $> |C^*|/2$, but $T$ also contains at least $s/2 = |C^*_i|/2$ points, some point in $T$ must not be in $C^*_i$, and therefore have distance larger than $\frac{3 \beta \text{OPT}}{4|C^*|}$. \ref{lemma:4.2b} then tells us all points must meet the property.

@@

We will show that $d(c_i, c_i*) < \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$ for a cheap cluster. We know by the definition of a cheap cluster that the cost is at most $(\epsilon/32)\beta\text{OPT}/|C_i^*|$. In order not exceed this bound, at least half of the points have distance at most $(\epsilon/16)\beta\text{OPT}/|C_i^*|$ from $c_i^*$. Also these are by definition in the inner ring. We know that at least half the points in a cheap cluster reside in the inner ring. If $d(c_i, c_i^*)  \geq \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$, then we require at least half of the points to contribute at least $(\epsilon/16)\beta\text{OPT}/|C_i^*|$ to the sum, which means some point in the inner ring must do so. Therefore $\sum\limits_{x\in T\cup B(T)}d(x, c_i)> OPT_i$, which contradicts the optimality of the inner ring points.

We therefore for each $x \in C_i^* and x\notin T\cup B(T)$ we have $d(c_i, c_i^*) \leq \epsilon\frac{\beta\text{OPT}}{8|C_i^*|}$. So $d(x, c_i) \leq d(x, c_i^*) + d(c_i^*, c_i) \leq (1+\epsilon)d(x,c_i^*)$. And so each point adds at most a factor of $(1+\epsilon)$ to the error, bounding our total error to $(1+\epsilon)\text{OPT}$
